---
output: github_document
editor_options: 
  chunk_output_type: console
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# beaver

<!-- badges: start -->
[![check-standard](https://github.com/rich-payne/beaver/actions/workflows/check-standard.yaml/badge.svg)](https://github.com/rich-payne/beaver/actions/workflows/check-standard.yaml)
[![test-coverage](https://github.com/rich-payne/beaver/actions/workflows/test-coverage.yaml/badge.svg)](https://github.com/rich-payne/beaver/actions/workflows/test-coverage.yaml)
[![lint](https://github.com/rich-payne/beaver/actions/workflows/lint.yaml/badge.svg)](https://github.com/rich-payne/beaver/actions/workflows/lint.yaml)
<!-- badges: end -->

The goal of beaver is to fit Bayesian model averaging of negative-binomial dose-response models.

We begin with an example where we simulate negative binomial data where age and gender are prognostic factors.  In this example, males have higher counts than females and older individual have higher counts than younger people.

```{r}
library(dplyr)
library(ggplot2)
library(beaver)
set.seed(222)
n <- 200
x <- data.frame(
  age = log(runif(n, 18, 65)),
  gender = factor(sample(c("F", "M"), n, replace = TRUE))
) %>%
  model.matrix(~age + gender, data = .)
df <- data_negbin_emax(
  n_per_arm = 50,
  doses = 0:3,
  b1 = c(-2, .75, .5),
  # b1 = c(-1, 0, 0),
  b2 = -2,
  b3 = 1.5,
  ps = .5,
  x = x
) %>%
  mutate(
    gender = case_when(
      genderM == 1 ~ "M",
      TRUE ~ "F"
    ),
    gender = factor(gender)
  ) %>%
  dplyr::select(subject, dose, age, gender, response)

data_sumry <- df %>%
  group_by(dose) %>%
  summarize(
    response = mean(response),
    age = mean(age),
    male = mean(gender == "M")
  )

ggplot(df, aes(dose, response, color = age)) +
  geom_point() +
  geom_jitter() +
  facet_grid(~ gender, labeller = label_both)
```

We now fit fit a Bayesian model where each dose is treated independently (no dose response) without covariates:

```{r}
mcmc_indep <- beaver_mcmc(
  indep = model_negbin_indep(
    mu_b1 = 0,
    sigma_b1 = 10,
    mu_b2 = 0,
    sigma_b2 = 10,
    w_prior = 1
  ),
  formula = ~ 1,
  data = df,
  n_adapt = 1e4,
  n_burn = 1e4,
  n_iter = 1e4,
  n_chains = 4,
  quiet = FALSE
)
# convergence
coda::gelman.diag(mcmc_indep$models$indep$mcmc, multivariate = FALSE)
# posterior mean at each dose
post_ind <- posterior(mcmc_indep, contrast = matrix(1, 1, 1))
post_ind$stats
# summary of data
data_sumry

plot(mcmc_indep, contrast = matrix(1, 1, 1))
```

We now fit fit a Bayesian model where each dose is treated independently (no dose response), but now include covariates:

```{r}
mcmc_cov_indep <- beaver_mcmc(
  indep = model_negbin_indep(
    mu_b1 = 0,
    sigma_b1 = 10,
    mu_b2 = 0,
    sigma_b2 = 10,
    w_prior = 1
  ),
  formula = ~ age + gender,
  data = df,
  n_adapt = 1e4,
  n_burn = 1e4,
  n_iter = 1e4,
  n_chains = 4,
  quiet = FALSE
)

coda::gelman.diag(mcmc_cov_indep$models$indep$mcmc, multivariate = FALSE)
# Bayesian g-computation estimate
post_cov_ind <- posterior_g_comp(mcmc_cov_indep, new_data = df)
# compare widths of covariate adjusted and non-covariate adjusted
mutate(post_cov_ind$stats, width = `97.50%` - `2.50%`)
mutate(post_ind$stats, width = `97.50%` - `2.50%`)
# data_sumry
plot(mcmc_cov_indep, new_data = df, type = "g-comp")
```

# Bayesian Model Averaging

We now fit Bayesian dose-respone models with and without covariate adjustment.

```{r}
mcmc <- beaver_mcmc(
  emax = model_negbin_emax(
    mu_b1 = 0,
    sigma_b1 = 10,
    mu_b2 = 0,
    sigma_b2 = 10,
    mu_b3 = 1.5,
    sigma_b3 = 3,
    w_prior = 1 / 7
  ),
  sigmoid_emax = model_negbin_sigmoid_emax(
    mu_b1 = 0,
    sigma_b1 = 10,
    mu_b2 = 0,
    sigma_b2 = 10,
    mu_b3 = 1.5,
    sigma_b3 = 3,
    mu_b4 = 1,
    sigma_b4 = 10,
    w_prior = 1 / 7
  ),
  linear = model_negbin_linear(
    mu_b1 = 0,
    sigma_b1 = 10,
    mu_b2 = 0,
    sigma_b2 = 10,
    w_prior = 1 / 7
  ),
  loglinear = model_negbin_loglinear(
    mu_b1 = 0,
    sigma_b1 = 10,
    mu_b2 = 0,
    sigma_b2 = 10,
    w_prior = 1 / 7
  ),
  quad = model_negbin_quad(
    mu_b1 = 0,
    sigma_b1 = 10,
    mu_b2 = 0,
    sigma_b2 = 10,
    mu_b3 = 1.5,
    sigma_b3 = 3,
    w_prior = 1 / 7
  ),
  logquad = model_negbin_logquad(
    mu_b1 = 0,
    sigma_b1 = 10,
    mu_b2 = 0,
    sigma_b2 = 10,
    mu_b3 = 1.5,
    sigma_b3 = 3,
    w_prior = 1 / 7
  ),
  exp = model_negbin_exp(
    mu_b1 = 0,
    sigma_b1 = 10,
    mu_b2 = 0,
    sigma_b2 = 10,
    mu_b3 = 0,
    sigma_b3 = 3,
    w_prior = 1 / 7
  ),
  formula = ~ 1,
  data = df,
  n_adapt = 1e4,
  n_burn = 1e4,
  n_iter = 1e4,
  n_chains = 4,
  quiet = FALSE
)

mcmc_cov <- beaver_mcmc(
  emax = model_negbin_emax(
    mu_b1 = 0,
    sigma_b1 = 10,
    mu_b2 = 0,
    sigma_b2 = 10,
    mu_b3 = 1.5,
    sigma_b3 = 3,
    w_prior = 1 / 7
  ),
  sigmoid_emax = model_negbin_sigmoid_emax(
    mu_b1 = 0,
    sigma_b1 = 10,
    mu_b2 = 0,
    sigma_b2 = 10,
    mu_b3 = 1.5,
    sigma_b3 = 3,
    mu_b4 = 1,
    sigma_b4 = 10,
    w_prior = 1 / 7
  ),
  linear = model_negbin_linear(
    mu_b1 = 0,
    sigma_b1 = 10,
    mu_b2 = 0,
    sigma_b2 = 10,
    w_prior = 1 / 7
  ),
  loglinear = model_negbin_loglinear(
    mu_b1 = 0,
    sigma_b1 = 10,
    mu_b2 = 0,
    sigma_b2 = 10,
    w_prior = 1 / 7
  ),
  quad = model_negbin_quad(
    mu_b1 = 0,
    sigma_b1 = 10,
    mu_b2 = 0,
    sigma_b2 = 10,
    mu_b3 = 1.5,
    sigma_b3 = 3,
    w_prior = 1 / 7
  ),
  logquad = model_negbin_logquad(
    mu_b1 = 0,
    sigma_b1 = 10,
    mu_b2 = 0,
    sigma_b2 = 10,
    mu_b3 = 1.5,
    sigma_b3 = 3,
    w_prior = 1 / 7
  ),
  exp = model_negbin_exp(
    mu_b1 = 0,
    sigma_b1 = 10,
    mu_b2 = 0,
    sigma_b2 = 10,
    mu_b3 = 0,
    sigma_b3 = 3,
    w_prior = 1 / 7
  ),
  formula = ~ age + gender,
  data = df,
  n_adapt = 1e4,
  n_burn = 1e4,
  n_iter = 1e4,
  n_chains = 4,
  quiet = FALSE
)

# diagnostics for each model
coda::gelman.diag(mcmc$models$exp$mcmc)
coda::gelman.diag(mcmc$models$emax$mcmc)
coda::gelman.diag(mcmc$models$sigmoid_emax$mcmc)
coda::gelman.diag(mcmc$models$linear$mcmc)
coda::gelman.diag(mcmc$models$quad$mcmc)
coda::gelman.diag(mcmc$models$logquad$mcmc)
coda::gelman.diag(mcmc$models$loglinear$mcmc)
# posterior weight
mcmc$w_post
# posterior estimate
post <- posterior(mcmc, contrast = matrix(1, 1, 1), doses = 0:3)
post$stats

# covariate adusted
coda::gelman.diag(mcmc_cov$models$exp$mcmc)
coda::gelman.diag(mcmc_cov$models$emax$mcmc)
coda::gelman.diag(mcmc_cov$models$sigmoid_emax$mcmc)
coda::gelman.diag(mcmc_cov$models$linear$mcmc)
coda::gelman.diag(mcmc_cov$models$quad$mcmc)
coda::gelman.diag(mcmc_cov$models$loglinear$mcmc)
coda::gelman.diag(mcmc_cov$models$logquad$mcmc)

post_cov_g <- posterior_g_comp(mcmc_cov, new_data = df)
post_cov_g$stats

# no covariate adjustment
plot(mcmc, contrast = matrix(1, 1, 1))
# with covariate adjustment
plot(mcmc_cov, new_data = df, type = "g-comp")
```

The following plot shows the widths of the credible intervals comparing the covariate-adjusted and unadjusted analyses.

```{r}
# compare widths
w_indep <- mutate(
  post_ind$stats,
  width = `97.50%` - `2.50%`,
  model = "indep",
  covariates = FALSE
) %>%
  select(dose, model, covariates, mean = value, `2.50%`, `97.50%`, width)

w_indep_cov <- mutate(
  post_cov_ind$stats,
  width = `97.50%` - `2.50%`,
  model = "indep",
  covariates = TRUE
) %>%
  select(dose, model, covariates, mean = value, `2.50%`, `97.50%`, width)

w_bma <- mutate(
  post$stats,
  width = `97.50%` - `2.50%`,
  model = "bma",
  covariates = FALSE
) %>%
  select(dose, model, covariates, mean = value, `2.50%`, `97.50%`, width)

w_bma_cov <- mutate(
  post_cov_g$stats,
  width = `97.50%` - `2.50%`,
  model = "bma",
  covariates = TRUE
) %>%
  select(dose, model, covariates, mean = value, `2.50%`, `97.50%`, width)

widths <- bind_rows(w_indep, w_indep_cov, w_bma, w_bma_cov)

ggplot(widths, aes(model, width, fill = covariates)) +
  geom_bar(stat = "identity", position = "dodge") +
  facet_wrap(~dose, labeller = label_both)
```

# Installation

```{r, eval = FALSE}
install.packages("beaver")
```
